Main Analyses of the Combined Dataset

### 2021-08-08 manuscript/elife/elife-manuscript.tex
updated figure 3 supp

  \figsupp[The fraction of total species on earth included in this study's
  sample, per class.]{The color of the points represents phylum, and the size of
the point represents the absolute number of species by
class.}{\includegraphics{species_counts.pdf}}\label{figsupp:species-counts}

\figsupp[Validation of this paper's range estimates procedure against the IUCN Red List's range estimates.]{
  The correspondence between the ranges estimated with the alpha hull
  method applied to GBIF data used in this paper and IUCN Red List's Extent
  of Occurrence for the subset of species in both datasets. Note that the
  IUCN Red List contains predominantly endangered species, which leads to
  ascertainment bias; still, the high correlation between the estimated
ranges shows the alpha hull method works well.}{\includegraphics[]{iucn_redlist_ranges.pdf}}\label{figsupp:redlist-ranges}

\figsupp[Validation of this paper's range estimates against the categorical labels of \cite{Leffler2012-zj}.]{
  The estimated ranges using GBIF occurrence data, ordered within and
  colored by the original range category labels assigned in
\cite{Leffler2012-zj}.}{\includegraphics[]{range_categories.pdf}}\label{figsupp:range-cats}

@@

A major impediment in quantifying Lewontin's Paradox has been estimating census
population sizes across many taxa, especially for extremely abundant,
cosmopolitan species that define the upper limit of ranges. Previous work has
surveyed the literature for census size estimates
\citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
qualitative categories as proxies for census size
\citep{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). My
approach predicts population densities from body sizes using a
previously-observed linear relationship that holds across metazoans (see
\FIGSUPP[figure-nc]{damuth}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
geographic occurrence data, I estimate range sizes.  Finally, I estimate
population size as the product of these predicted densities and range estimates
(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the relationship between
population density and body size is driven by energy budgets, and thus reflects
macroecological equilibria \citep{Damuth1987-sg}. Consequently, population
sizes are underestimated for taxa like humans and their domesticated species,
and overestimated for species with anthropogenically reduced densities or
fragmented ranges. For example, the population size of \emph{Lynx lynx} is
likely around 50,000 \citep{Iucn2020-ap} which is around two orders of
magnitude smaller than my estimate. Additionally, the range size estimates do
not consider whether an area has unsuitable habitat, and thus may be
overestimated for species with particular niches or patchy habitats. While
these methods to estimate census size are crude and approximate, they can be
efficiently calculated for numerous taxa and are sufficient to estimate the
scale of Lewontin's Paradox (see \nameref{supp:popsize-valid} for more on
validation based on biomass and other approaches).

\subsection{Quantifying Lewontin's Paradox}

@@

  \caption{A visualization of Lewontin's Paradox of Variation. Pairwise
    diversity (data from \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt},
    and \cite{Romiguier2014-bp}), which varies around three orders of
    magnitude, shows a weak relationship with approximate population size,
    which varies over 12 orders of magnitude.  The shaded curve shows the range
    of expected neutral diversity if $N_e$ were to equal $N_c$ under the
    four-alleles model, $\log_{10}(\pi) = \logt(\theta) - \logt{(1 +
    \nicefrac{4\theta}{3})}$ where $\theta = 4N_c \mu$, for two mutation rates,
    $\mu = 10^{-8}$ and $\mu = 10^{-9}$, and the light gray dashed line
    represents the maximum pairwise diversity under the four alleles model. The
    dark gray dashed line is the OLS regression fit, and the blue dashed line
    is the regression fit using a phylogenetic mixed-effects model.  Points are
    colored by phylum. The species \emph{Equus ferus przewalskii} ($N_c \approx
    10^3$ and $\pi = 3.6 \times 10^{-3}$) was an outlier and excluded from this
    figure for visual clarity.}\label{fig:pi-Nc}

@@

To determine which ecological or evolutionary processes could decouple
diversity from census population size, we first need to quantify this
relationship across a wide variety of taxa. Previous work has found there is a
significant relationship between heterozygosity and the logarithm of population
size or range size, but these studies relied on heterozygosity measured from
allozyme data \citep{Soule1976-he,Frankham1996-yb,Nei1984-zi,Leffler2012-zj}.
Here, I confirm these findings using pairwise diversity estimates from genomic
sequence data and the estimated census sizes (\FIG{pi-Nc}).  The pairwise
diversity estimates are from three sources: \cite{Leffler2012-zj},
\cite{Corbett-Detig2015-gt}, and \cite{Romiguier2014-bp}, and are predominantly
from either synonymous or non-coding DNA (see Methods and Materials:
\nameref{sec:methods-diversity}). Overall, an ordinary least squares (OLS)
relationship on a log-log scale fits the data well (\FIG{pi-Nc}, gray dashed
line). The OLS slope estimate is significant and implies a 13\% percent
increase in differences per basepair for every order of magnitude census size
grows (95\% confidence interval $[12\%, 14\%]$, adjusted $R^2 = 0.26$; see also
the OLS fit per-phyla, \FIGSUPP[pi-Nc]{popsize-ave}).

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\citep{Nei1984-zi,Soule1976-he}; see \FIGSUPP[pi-Nc]{popsize-linear} for a
version on a log-linear scale. However, it is noteworthy that few taxa have
diversity estimates below $10^{-3.5}$ differences per basepair. Those that do,
lynx (\emph{Lynx Lynx}), wolverine (\emph{Gulo gulo}), and Massasauga
rattlesnake (\emph{Sistrurus catenatus}) face habitat fragmentation and
declining population sizes. These three species are all in the IUCN Red List,
but are listed as least concern (though their presence in the Red List
indicates they are of conservation interest). In Appendix \ref{app:div-iucn},
\nameref{app:div-iucn}, I explore more about the relationships between IUCN Red
List status, diversity, and population size.

@@

One limitation of using ordinary least squares is that shared phylogenetic
history can create correlation structure in the residuals, which violates an
assumption of the regression model
\citep{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I
fit this relationship using a phylogenetic mixed-effects model, investigated
whether there is a signal of phylogenetic non-independence, estimated the
continuous trait values on the phylogeny, and explored how diversity and
population size evolve. Prior population genetic comparative studies have
lacked time-calibrated phylogenies and assumed unit branch lengths
\citep{Whitney2010-ud}, a shortcoming that has drawn criticism
\citep{Lynch2011-qv}. Here, I use a synthetic time-calibrated phylogeny
created from the DateLife project \citep{OMeara2020-ds} to account for
shared phylogenetic history (see Methods and Materials: \nameref{sec:methods-pcm}).

@@

Using a phylogenetic mixed-effects model
\citep{Lynch1991-tp,Hadfield2010-ql,De_Villemereuil2014-kt} implemented in Stan
\citep{Carpenter2017-xi,Stan_Development_Team2020-ea}, I estimated the linear
relationship between diversity and population size (on a log-log scale)
accounting for phylogeny, for the 166 taxa without missing data and present in
the synthetic chronogram. As with the linear regression, this relationship was
positive and significant (95\% credible interval $0.03, 0.11$), though somewhat
attenuated compared to the OLS estimates (\FIG{pcm}B). Since
the population size estimates are based on range and body mass, they are
essentially a composite trait; fitting phylogenetic mixed-effects models
separately on body mass and range indicates these have significant negative and
positive effects, respectively (\FIGSUPP[pi-Nc]{div-range-bodymass}; see also
\FIGSUPP[pi-Nc]{div-leffler} for the relationship between diversity and the
range categories of \cite{Leffler2012-zj}).

With the phylogenetic mixed-effects model, I also estimated the variance of the
phylogenetic effect ($\sigma_p^2$) and the residual variance ($\sigma_r^2$),
which can be used to estimate a measure of the phylogenetic signal, $\lambda =
\nicefrac{\sigma_p^2}{(\sigma_p^2 + \sigma_r^2)}$
(\cite{Lynch1991-tp,De_Villemereuil2014-kt}; see \cite{Freckleton2002-ly} for a
comparison to Pagel's $\lambda$).  If the relationship between diversity and
population size was free of shared phylogenetic history, $\lambda = 0$ and all
the variance could be explained by evolution on the tips; this is analogous to
Lynch's conjecture that coalescent times should be free of phylogenetic signal
(\citeyear{Lynch2011-qv}). In the relationship between population size and
diversity, the posterior mean of $\lambda = 0.67$ (90\% credible interval
$[0.58, 0.75]$) indicates that the majority of the variance perhaps might be
due to shared phylogenetic history (\FIG{pcm}B).

@@

To investigate whether clade-level differences dominated the relationship
between diversity and population size, I fit phylogenetic mixed-effects models
to phyla-level subsets of the data for clades with sufficient sample sizes (see
Methods: \nameref{sec:methods-pcm}). This analysis shows a significant positive
relationship between diversity and population size in arthropods, and positive
weak relationships in molluscs and chordates
(\FIGSUPP[pcm]{div-popsize-post-phyla}). Each of the 90\% credible intervals
for slope overlap, indicating the relationship between $\pi$ and $N_c$ is
similar across these clades.

@@

Interestingly, the diversity node-height test reveals two rate shifts at deeper
splits (\FIG{pcm}C, top left) around 570 Mya. These nodes
represent the branches between tunicates and vertebrates in chordates, and
cephalopods and pleistomollusca (bivalves and gastropods) in molluscs.  While
the cephalopod-pleistomollusca split outlier may be an artifact of having a
single cephalopod (\emph{Sepia officinalis}) in the phylogeny, the
tunicate-vertebrate split outlier is driven by the low diversity of vertebrates
and the previously-documented exceptionally high diversity of tunicates (sea
squirts; \cite{Nydam2010-kg,Small2007-mt}). This deep node representing a rate
shift in diversity could reflect a change in either effective population size
or mutation rate, and there is some evidence of both in this genus \emph{Ciona}
\citep{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts
in diversity is mirrored in the population size node-height test (Figure
\FIG{pcm}C, top right). Rather, it appears a trait impacting diversity
but not census size (e.g. mutation rate or offspring distributions) has
experienced a shift on the lineage separating tunicates and vertebrates. At
nearly 600 Mya, these deep nodes illustrate that expected coalescent times can
share phylogenetic history, due to phylogenetic inertia in some combination of
population size, reproductive system, and mutation rates.

Finally, an important caveat is the increase in rate towards the tips could be
caused by measurement noise, or possibly uncertainty or bias in the divergence
time estimates deep in the tree. Inspecting the lineage pairs that lead to this
increase in rate towards the tips indicates these represent plausible rate
shifts, e.g.  between cosmopolitan and endemic sister species like
\emph{Drosophila simulans} and \emph{Drosophila sechellia}; however, ruling out
measurement noise entirely as an explanation would involve considering the
uncertainty of diversity and population size estimates.

@@

The above analyses reemphasize the drastic shortfall of diversity levels as
compared to census sizes. Linked selection has been proposed as the mechanism
that acts to reduce diversity levels from what we would expect given census
sizes \citep{Maynard_Smith1974-zr,Gillespie2000-mh,Corbett-Detig2015-gt}.
Here, I test this hypothesis by estimating the scale of diversity reductions
expected under background selection and recurrent hitchhiking, and compare
these to the observed relationship between $\pi$ and $N_c$.

I quantify the effect of linked selection on diversity as the ratio of observed
diversity ($\pi$) to the estimated diversity in the absence of linked selection
($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. Here, $\pi_0$ would reflect only
demographic history and non-heritable variation in reproductive success. There
are two difficulties in evaluating whether linked selection could resolve
Lewontin's Paradox. The first difficulty is that $\pi_0$ is unobserved.
Previous work has estimated $\pi_0$ using methods that exploit the spatial
heterogeneity in recombination and functional density across the genome to fit
linked selection models that incorporate both hitchhiking and background
selection \citep{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
difficulty is understanding of how $R$ varies across taxa, since we lack
estimates of critical model parameters for most species. Still, I can address a
key question: if diversity levels were determined by census sizes ($\pi_0 =
4N_c \mu$), are the combined effects of background selection and recurrent
hitchhiking sufficient to reduce diversity to observed levels? Furthermore,
does the relationship between census size and predicted diversity under linked
selection across species, $\pi_{BGS+HH} = R \pi_0$, match the observed
relationship in \FIG{pi-Nc}?

@@

>>>>

  \figsupp[The fraction of total species per class on earth included in this study's
  sample, per class.]{The color of the points represents phylum, and the size of
the point represents the absolute number of species by
class.}{\includegraphics{species_counts.pdf}}\label{figsupp:species-counts}

\figsupp[Comparison of this paper's range estimates procedure against the IUCN Red List's range estimates.]{
  The correspondence between the ranges estimated with the alpha hull
  method applied to GBIF data used in this paper and IUCN Red List's Extent
  of Occurrence for the subset of species in both datasets. Note that the
  IUCN Red List contains predominantly endangered species, which leads to
  ascertainment bias; still, the high correlation between the estimated
ranges shows the alpha hull method works well.}{\includegraphics[]{iucn_redlist_ranges.pdf}}\label{figsupp:redlist-ranges}

\figsupp[Validation of this paper's range estimates against the categorical labels of \cite{Leffler2012-zj}.]{
  The estimated ranges using GBIF occurrence data, ordered within and
  c olored by the original range category labels assigned in
\cite{Leffler2012-zj}.}{\includegraphics[]{range_categories.pdf}}\label{figsupp:range-cats}

@@

An impediment in resolving Lewontin's Paradox is characterizing the
relationship between diversity and census population sizes. This is difficult
because census population sizes are unavailable for many taxa, especially for
extremely abundant, cosmopolitan species that define the upper limit of ranges.
Previous work has surveyed the literature for census size estimates
\citep{Nei1984-zi,Soule1976-he,Frankham1996-yb}, or used range, body size, or
qualitative categories as proxies for census size
\citep{Corbett-Detig2015-gt,Leffler2012-zj}. To quantify the relationship
between genomic estimates of diversity and census population sizes, I first
approximate census population sizes for 172 metazoan taxa (\FIG{figure-nc}). I
estimate population densities based on an empirical linear relationship between
body sizes and density that holds across metazoans (see
\FIGSUPP[figure-nc]{damuth}; \cite{Damuth1981-st,Damuth1987-sg}). Then, from
geographic occurrence data, I estimate range sizes.  Finally, I estimate
population size as the product of these predicted densities and range estimates
(see Methods and Materials: \nameref{sec:methods-popsize}). Note that the
relationship between population density and body size is driven by energy
budgets, and thus reflects macroecological equilibria \citep{Damuth1987-sg}.
Consequently, population sizes are underestimated for taxa like humans and
their domesticated species, and overestimated for species with
anthropogenically reduced densities or fragmented ranges. For example, the
population size of \emph{Lynx lynx} is likely around 50,000 \citep{Iucn2020-ap}
which is around two orders of magnitude smaller than my estimate. Additionally,
the range size estimates do not consider whether an area has unsuitable
habitat, and thus may be overestimated for species with particular niches or
patchy habitats. While my approach produces approximate and sometimes crude
estimates, it has the advantage that it can be efficiently calculated for
numerous taxa, which is sufficient to estimate the magnitude of Lewontin's
Paradox (see \nameref{supp:popsize-valid} for more on validation based on
biomass and other approaches).

\subsection{Characterizing the Diversity--Census-size Relationship}

@@

  \caption{A visualization of Lewontin's Paradox of Variation. Pairwise
    diversity (data from \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt},
    and \cite{Romiguier2014-bp}), which varies over three orders of
    magnitude, shows a weak relationship with approximate population size,
    which varies over 12 orders of magnitude.  The shaded curve shows the range
    of expected neutral diversity if $N_e$ were to equal $N_c$ under the
    four-alleles model, $\log_{10}(\pi) = \logt(\theta) - \logt{(1 +
    \nicefrac{4\theta}{3})}$ where $\theta = 4N_c \mu$, for two mutation rates,
    $\mu = 10^{-8}$ and $\mu = 10^{-9}$, and the light gray dashed line
    represents the maximum pairwise diversity under the four alleles model. The
    dark gray dashed line is the OLS regression fit, and the blue dashed line
    is the regression fit using a phylogenetic mixed-effects model.  Points are
    colored by phylum. The species \emph{Equus ferus przewalskii} ($N_c \approx
    10^3$ and $\pi = 3.6 \times 10^{-3}$) was an outlier and excluded from this
    figure for visual clarity.}\label{fig:pi-Nc}

@@

To determine which ecological or evolutionary processes could decouple
diversity from census population size, we first need to quantify this
relationship across a wide variety of taxa. Previous work has found there is a
significant relationship between heterozygosity and the logarithm of population
size or range size, but these studies relied on heterozygosity measured from
allozyme data \citep{Soule1976-he,Frankham1996-yb,Nei1984-zi}. I confirm these
findings using pairwise diversity estimates from genomic sequence data and the
estimated census sizes (\FIG{pi-Nc}). The pairwise diversity estimates are from
three sources: \cite{Leffler2012-zj}, \cite{Corbett-Detig2015-gt}, and
\cite{Romiguier2014-bp}, and are predominantly from either synonymous or
non-coding DNA (see Methods and Materials: \nameref{sec:methods-diversity}).
Overall, an ordinary least squares (OLS) relationship on a log-log scale fits
the data well (\FIG{pi-Nc}, gray dashed line). The OLS slope estimate is
significant and implies a 13\% percent increase in differences per basepair for
every order of magnitude census size grows (95\% confidence interval $[12\%,
14\%]$, adjusted $R^2 = 0.26$; see also the OLS fit per-phyla,
\FIGSUPP[pi-Nc]{popsize-ave}).

Notably, this relationship has few outliers and is relatively homoscedastic.
This is in part because of the log-log scale, in contrast to previous work
\citep{Nei1984-zi,Soule1976-he}; see \FIGSUPP[pi-Nc]{popsize-linear} for a
version on a log-linear scale. However, it is noteworthy that few taxa have
diversity estimates below $10^{-3.5}$ differences per basepair. Those that do,
lynx (\emph{Lynx Lynx}), wolverine (\emph{Gulo gulo}), and Massasauga
rattlesnake (\emph{Sistrurus catenatus}) face habitat loss and declining
population sizes. These three species are all in the IUCN Red List, but are
listed as least concern (though their presence in the Red List indicates they
are of conservation interest). In Appendix \ref{app:div-iucn},
\nameref{app:div-iucn}, I explore the relationships between IUCN Red List
status, diversity, and population size.

@@

One limitation of using ordinary least squares is that shared phylogenetic
history can create correlation structure in the residuals, which violates an
assumption of the regression model and can lead to bias
\citep{Felsenstein1985-an,Revell2010-mf}. To address this shortcoming, I fit
the diversity--census-size relationship using a phylogenetic mixed-effects
model, investigated whether there is a signal of phylogenetic non-independence,
estimated the continuous trait values on the phylogeny, and explored how
diversity and population size evolve. Prior population genetic comparative
studies have lacked time-calibrated phylogenies and assumed unit branch lengths
\citep{Whitney2010-ud}, a shortcoming that has drawn criticism
\citep{Lynch2011-qv}.  I use a synthetic time-calibrated phylogeny created from
the DateLife project \citep{OMeara2020-ds} to account for shared phylogenetic
history (see Methods and Materials: \nameref{sec:methods-pcm}).

@@

Using a phylogenetic mixed-effects model
\citep{Lynch1991-tp,Hadfield2010-ql,De_Villemereuil2014-kt} implemented in Stan
\citep{Carpenter2017-xi,Stan_Development_Team2020-ea}, I estimated the linear
relationship between diversity and population size (on a log-log scale)
accounting for phylogeny, for the 166 taxa without missing data and present in
the synthetic chronogram. This phylogenetic mixed-effects model can account for
correlated residuals among more closely related species when estimating the
relationship between $N_c$ and $\pi$, but does not assume there is phylogenetic
signal in either of these variables. Since the phylogenetic mixed-effects model
simultaneously estimates the degree of phylogenetic structure in the residuals
while fitting the relationship between $N_c$ and $\pi$, this model's estimates
would match those found by ordinary least squares if the residuals were
distributed independently. Overall, this approach is conservative, making no
assumptions about the source of the phylogenetic signal while accounting for
violations of the regression model due to dependence among the residuals (see
\cite{Revell2010-mf} for a discussion of this).

As with the linear regression, I find this relationship is positive and
significant (95\% credible interval $0.03, 0.11$), though somewhat attenuated
compared to the OLS estimates (\FIG{pcm}B). Since the population size estimates
are based on range and body mass, they are essentially a composite trait;
fitting phylogenetic mixed-effects models separately on body mass and range
indicates these have significant negative and positive effects, respectively
(\FIGSUPP[pi-Nc]{div-range-bodymass}; see also \FIGSUPP[pi-Nc]{div-leffler} for
the relationship between diversity and the range categories of
\cite{Leffler2012-zj}).

Since the phylogenetic mixed-effects model simultaneously estimates the
variance of the phylogenetic effect ($\sigma_p^2$) and the residual variance
($\sigma_r^2$), these can be used to estimate the phylogenetic signal, $\lambda
= \nicefrac{\sigma_p^2}{(\sigma_p^2 + \sigma_r^2)}$
(\cite{Lynch1991-tp,De_Villemereuil2014-kt}; see \cite{Freckleton2002-ly} for a
comparison to Pagel's $\lambda$). If the residuals were free of correlations
due to shared phylogenetic history, then $\lambda = 0$ and all the variance
could be explained by evolution on the tips; this is analogous to Lynch's
conjecture that coalescent times should be free of phylogenetic signal
(\citeyear{Lynch2011-qv}). In the relationship between population size and
diversity, the posterior mean of $\lambda = 0.67$ (90\% credible interval
$[0.58, 0.75]$) indicates a majority of the variance perhaps might be
due to shared phylogenetic history (\FIG{pcm}B).

@@

To investigate whether clade-level differences dominated the relationship
between diversity and population size, I fit phylogenetic mixed-effects models
to phyla-level subsets of the data for clades with sufficient sample sizes (see
Methods: \nameref{sec:methods-pcm}). This analysis shows a significant positive
relationship between diversity and population size in arthropods, and positive
weak relationships in molluscs and chordates
(\FIGSUPP[pcm]{div-popsize-post-phyla}). Each of the 90\% credible intervals
for slope overlap, suggesting the relationship between $\pi$ and $N_c$ is
similar across these clades.

@@

Interestingly, the diversity node-height test reveals two rate shifts at deeper
splits (\FIG{pcm}C, top left) around 570 Mya. These nodes represent the
branches between tunicates and vertebrates in chordates, and cephalopods and
pleistomollusca (bivalves and gastropods) in molluscs.  While the
cephalopod-pleistomollusca split outlier may be an artifact of having a single
cephalopod (\emph{Sepia officinalis}) in the phylogeny, the tunicate-vertebrate
split outlier is driven by the low diversity of vertebrates and the
previously-documented exceptionally high diversity of tunicates (sea squirts;
\cite{Nydam2010-kg,Small2007-mt}). This deep node representing a rate shift in
diversity could reflect a change in either effective population size or
mutation rate, and there is some evidence of both in this genus \emph{Ciona}
\citep{Small2007-mt,Tsagkogeorga2012-lv}. Neither of these deep rate shifts in
diversity is mirrored in the population size node-height test (Figure
\FIG{pcm}C, top right). Rather, it appears a trait impacting diversity but not
census size (e.g. mutation rate or offspring distributions) has experienced a
shift on the lineage separating tunicates and vertebrates. At nearly 600 Mya,
these deep nodes illustrate that expected effective population sizes (and thus
coalescent times) can share phylogenetic history, due to phylogenetic inertia
in some combination of population size, reproductive system, and mutation
rates.

Finally, an important caveat is the increase in rate towards the tips could be
caused by measurement noise, or possibly that uncertainty or bias in the
divergence time estimates deep in the tree. Inspecting the lineage pairs that
lead to this increase in rate towards the tips indicates these represent
plausible rate shifts, e.g.  between cosmopolitan and endemic sister species
like \emph{Drosophila simulans} and \emph{Drosophila sechellia}; however,
ruling out measurement noise entirely as an explanation would involve
considering the uncertainty of diversity and population size estimates.

@@

The above analyses reemphasize the drastic shortfall of diversity levels as
compared to census sizes. Linked selection has been proposed as the mechanism
that acts to reduce diversity levels from what we would expect given census
sizes \citep{Maynard_Smith1974-zr,Gillespie2000-mh,Corbett-Detig2015-gt}.
Here, I test this hypothesis by estimating the scale of diversity reductions
expected under background selection and recurrent hitchhiking, and comparing
these to the observed relationship between $\pi$ and $N_c$.

I quantify the effect of linked selection on diversity as the ratio of observed
diversity ($\pi$) to the estimated diversity in the absence of linked selection
($\pi_0$), $R = \nicefrac{\pi}{\pi_0}$. Here, $\pi_0$ would reflect only
demographic history and non-heritable variation in reproductive success. There
are two difficulties in evaluating whether linked selection could resolve
Lewontin's Paradox. The first difficulty is that $\pi_0$ is unobserved.
Previous work has estimated $\pi_0$ using methods that exploit the spatial
heterogeneity in recombination and functional density across the genome to fit
linked selection models that incorporate both hitchhiking and background
selection \citep{Elyashiv2016-vt,Corbett-Detig2015-gt}. The second
difficulty is understanding how $R$ varies across taxa, since we lack
estimates of critical model parameters for most species. Still, I can address a
key question: if diversity levels were determined by census sizes ($\pi_0 =
4N_c \mu$), are the combined effects of background selection and recurrent
hitchhiking sufficient to reduce diversity to observed levels? Furthermore,
does the relationship between census size and predicted diversity under linked
selection across species, $\pi_{BGS+HH} = R \pi_0$, match the observed
relationship in \FIG{pi-Nc}?

@@

  \figsupp{The observed $\pi$--$N_c$ relationship (points) across species
    compared to the predicted diversity (ribbons) under different modes of
    linked selection and parameters, for a range of mutation rates $\mu =$
    $10^{-8}--10^{-9}$. In both subplots, the gray ribbon is the expected
    diversity if $N_e = N_c$. In (A), the predicted impact on diversity for
    four modes of linked selection are depicted: background selection (purple)
    and hitchhiking (yellow) individually under the parameters in the
    \FIG{linksel}B, and strong background selection (red) where
    $U_\text{strongBGS} = 10 U_\text{Dmel} \approx 16$, and strong recurrent
    hitchhiking, where $\gamma_\text{strongHH} = 10 \gamma_\text{Dmel} \approx
    0.23$. (B) The combined strong background selection and strong hitchhiking
    (orange) compared to the original predicted diversity as in \FIG{linksel}B
(blue).}{\includegraphics[width=\textwidth]{figure_3_alt.pdf}}\label{figsupp:linked-sel-alt}

### 2021-08-08 notebooks/main_analysis.Rmd
revised ms
updated figure 3 supp

@@

@@

# this is the one used in figures -- assumes Ne = Nc
# this got a bit unwieldy as I tried different things... oh well.
dml_full <- da_ml %>%
     # estimate from Elyashiv et al.
     # mutate(J22 = 100*J22s(dm_seldist$s,  dm_seldist$g, 10^log10_popsize, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
     # RHH for comparison
     mutate(Ne_N_RHH = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=1, gamma=dm_gamma, 
                                   J22=J22)) %>%
     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=Ne_N_HK95, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%

     mutate(Ne_N_RHH_BGS_strongsel_strongBGS = RHH_BGS(1, L=map_length, 
                                                       N=10^log10_popsize,
                                             B=Ne_N_HK95_strongBGS, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%
     # 10-folder greater Nc
     mutate(Ne_N_RHH_BGS2 = RHH_BGS(1, L=map_length, N=10*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%

>>>>

What percent increase in pi is that for each order of magnitude change in N?

```{r}
signif(100 * 10^(coef(dps_ols)[2] - 1), 2)
ols_cis <- confint(dps_ols)[2, ]
signif(100 * 10^(ols_cis - 1), 2)
```

@@

Significance?

```{r}
sapply(list(popsize=psnh, range=rnnh, bodymass=bmnh, diversity=dvnh), function(x) x$pval)
```

@@

# this is the one used in figures -- assumes Ne = Nc
# this got a bit unwieldy as I tried different things... oh well.
dml_full <- da_ml %>%
     # estimate from Elyashiv et al.
     # mutate(J22 = 100*J22s(dm_seldist$s,  dm_seldist$g, 10^log10_popsize, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
     # RHH for comparison
     mutate(Ne_N_RHH = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=1, gamma=dm_gamma, 
                                   J22=J22)) %>%
     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=Ne_N_HK95, gamma=10*dm_gamma, 
                                             J22=J22)) %>%
     mutate(Ne_N_RHH_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=1, gamma=10*dm_gamma, 
                                             J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel_strongBGS = RHH_BGS(1, L=map_length, 
                                                       N=10^log10_popsize,
                                             B=Ne_N_HK95_strongBGS, gamma=10*dm_gamma, 
                                             J22=J22)) %>%
     # 10-folder greater Nc
     mutate(Ne_N_RHH_BGS2 = RHH_BGS(1, L=map_length, N=10*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
